Verifying Demonstrations

### 2020-03-06 paper/DB-Schnorr-UWB.tex
Updates UWB description to use CL04 for AKE

We will now introduce an extended protocol, \(\DBSHW\), which represents the 
secure hardware implementation.
\Cref{DBSHW-overview} presents an overview of our protocol.
For simplicity of exposition, we present a \ac{PKI} based on plain 
CL04~\cite{CLsignatures} signatures.
However, CL04 is the base for the \ac{DAA} protocol, which is used for 
\ac{TPM}, so our approach can make \ac{DAA} \acl{DB}.

In summary, the difference is as follows.
\(\DBS\) has a shared key~\(k\) and performs \(
  \PK[\alpha][A = g^\alpha]
\) with a distance bound.
\(\DBSHW\) first runs \iac{DHKE} to establish the shared key~\(k =
\hat g^{xy}\) from \(X = \hat g^x, Y = \hat g^y\), then performs \[
  \PK[\alpha, x, \sk_P, \sigma][
    A = g^\alpha \land
    X = \hat g^x \land
    \Blind[\sigma] = \Sign[_\sk][\sk_P]
  ],
\] also with a distance bound; where \(\sk_P\) is the prover's private key, 
\(\sigma\) is a signature issued with signing key~\(sk\) and \(\Blind\) 
reblinds a signature.
Proving knowledge of \(x\) prevents any \ac{MITM} attack against the \ac{DHKE}.
Proving knowledge of a signature~\(\sigma\) by \iac{CA} on the private 
key~\(\sk_P\) ensures that the prover is a device in the \ac{PKI} of the 
\ac{CA}.

\Cref{DBSHW-overview} provides an overview.

@@

    \multicolumn{2}{l}{%
      Static \(\sk_P\in \ZZ_q, \sigma = (a, A, b, B, c)\)
    }
    & Static \(\pk\)
    \\

@@

    \(x\rgets \ZZ_q\)
    & \(\ProtoSendRight{g^x}\)
    & \(y\rgets \ZZ_q\)
    \\

    \(k \gets (g^y)^x\)
    & \(\ProtoSendLeft{g^y}\)
    & \(k \gets (g^{x})^y\)
    \\

@@

    \(\rho, \hat\rho \rgets \ZZ_q, R\gets g^{\rho}, \hat R\gets g^{\hat\rho}\)
    & \(\ProtoSendRight{R, \hat R}\)
    & \(\hat A\gets g^x\)
    \\

    \(s_1\gets \rho - c_1\alpha \pmod q\)
    & \(\ProtoSendLeft{c_1, \dotsc, c_k}\)
    & \(c_1\rgets \bin^l, \dotsc, c_m\rgets \bin^l\)
    \\

    \(\hat s_1\gets \hat\rho - c_1 x \pmod q\)
    &
    &
    \\

    \(\hat s_m\gets \hat\rho - c_m x \pmod q\)
    &
    &
    \\

@@

    Fetch \(s_i, \hat s_i\)
    & \(\ProtoSendRight{s_i, \hat s_i}\)
    & Stop clock, record \(\Delta t\)
    \\

@@

    % null
    &
    & Accept if \(R = A^{c_i} g^{s_i}\),
      \(\hat R =  \hat A^{c_i} \hat g^{\hat s_i}\) and
      \(\Delta t < t_{\max}\).
    \\

  \end{tabular}
  \caption{%
    One-round protocol instance of the \(\DBSHW\) protocol instantiating 
    \(\PK[\alpha][A = g^\alpha]\).
    Each transmission (arrow in the diagram) uses \(\UWBPR\) (except in the 
    pre-setup phase).
    The protocol actually proves \(\PK[\alpha, x][A = g^\alpha \land
      \hat A = \hat g^x]\).
    This is ties \(x\), and consequently \(k\), to \(\alpha\).
    The protocol (setup, distance-bounding and verification phases) should be 
    repeated \(n\) times to achieve the desired soundness and distance-bounding 
    errors.
  }%
  \label{DBSHW-overview}
\end{figure*}

@@

We have public parameters \(q, G, g, \G, \g, e\); where \(e\colon G\times G\to 
\G\) is a bilinear map, \(g\in G\) and \(\g\in \G\) are generators of prime 
order \(q\) (hence \(e(g, g) = \g\)).

We let the \acg{CA} signing key be \(\sk = (x, y, z)\rgets \ZZ_q^3\) and the 
public key be \(\pk = (q, G, \G, g, \g, e, X, Y, Z)\), where \(X = g^x, Y = 
g^y, Z = g^z\).

The prover chooses \(\sk_P\rgets \ZZ_q\) and computes a commitment \(M = 
g^{\sk_P} Z^r\), where \(r\rgets \ZZ_q\).
Now the prover convinces the \ac{CA} that he knows the key by running the 
following \ac{ZKPK} protocol: \(
  \PK[\sk_P, r][M = g^{\sk_P} Z^r].
\)

Now the \ac{CA} computes the signature~\(\sigma = (a, A, b, B, c)\), where
\begin{itemize}
  \item \(a\gets g^\alpha\), where \(\alpha\rgets \ZZ_q\),
  \item \(A\gets g^z, b\gets a^y, B\gets A^y\),
  \item \(c\gets a^x M^{\alpha x y}\).
\end{itemize}
and gives to the prover.

@@

To establish \(k\), we use the \ac{DHKE}.
The prover chooses \(x\rgets \ZZ_q\), the verifier chooses \(y\rgets \ZZ_q\) 
and they exchange \(X = \hat g^x, Y = \hat g^y\) and both compute
\(k = \hat g^{xy}\), for some generator \(\hat g\) in \iac{DH} group of prime 
order \(q\).

We will see later that if the adversary attempted \iac{MITM} attack, our 
\(\DBSHW\) protocol will detect this, because the adversary will fail 
\iac{ZKPK}.

@@

The setup phase is similar to \(\DBS\) (\cref{SchnorrFigure}).
They use the agreed upon key~\(k = g^{xy}\) as the key for \(\UWBPR\).
\(\DBS\) performs \(
  \PK[\alpha][A = g^\alpha]
\) (with a distance bound) whereas \(\DBSHW\) must perform \(
  \PK[\alpha, x][\hat A = g^\alpha \land X = \hat g^x]
\) (also with a distance bound).

The prover chooses \(\rho, \hat\rho\rgets \ZZ_q\) and computes \(R\gets 
  g^\rho\) and \(\hat R\gets \hat g^{\hat\rho}\).
The verifier computes \(\hat A\gets \pk_P\).

The verifier chooses \(m\) challenges of length \(l\), \(c_1\rgets \bin^l, 
  \dotsc, c_m\gets \bin^l\), and sends these to the prover.
The prover computes responses~\(s_i\gets \rho - c_i\alpha\) and \(\hat s\gets 
  \hat\rho - c_i\sk_P\) for \(i\in \{1, \dotsc, m\}\).

@@

The verifier decides which of the \(m\) challenges to use, \(i\rgets \{1, 
  \dotsc, m\}\), sends its decision~\(i\) to the prover and starts its clock.
The prover instantly replies with the pre-computed \(s_i, \hat s_i\).
Note that \(s_i, \hat s_i\) must be sent as a concatenation in \emph{one} 
\(\UWBPR\) reply.
The verifier stops the clock and records the round-trip time~\(\Delta t\).

@@

The verifier checks if \(R = \hat A^{c_i} g^{s_i}\) and that \(\hat R =
  \hat g^{\hat s_i}\) and that the round-trip time \(\Delta t < t_{\max}\) does 
not exceed the limit~\(t_{\max}\).

>>>>

We will now introduce an extended protocol, \(\DBSHW\), which represents the 
secure hardware implementation.
\Cref{DBSHW-overview} presents an overview of our protocol.
For simplicity of exposition, we present a \ac{PKI} based on plain 
CL04~\cite{CLsignatures} signatures.
However, CL04 is the base for the \ac{DAA} protocol, which is used for 
\ac{TPM}, so our protocol can be modified into \iacl{DB} \ac{DAA} protocol.

In summary, the difference is as follows.
\(\DBS\) has a static shared key~\(k\) and performs \(
  \PK[\alpha][A = g^\alpha]
\) with a distance bound.
\(\DBSHW\) first runs \iac{DHKE} to establish the shared key~\(k =
\gDH^{\xDH \yDH}\) from \(\XDH = \gDH^{\xDH}, \YDH = \gDH^{\yDH}\), then 
performs \[
  \PK[\alpha, \xDH, \sk_P, \sigma][
    A = g^\alpha \land
    \XDH = \gDH^{\xDH} \land
    \Blind[\sigma] = \Sign[_{\sk}][\sk_P]
  ],
\] also with a distance bound; where \(\sk_P\) is the prover's private key, 
\(\sigma\) is a signature issued with signing key~\(\sk\) and \(\Blind\) 
reblinds a signature.
Proving knowledge of \(\xDH\) prevents any \ac{MITM} attack against the 
\ac{DHKE}.
Proving knowledge of a signature~\(\sigma\) by \iac{CA} on the private 
key~\(\sk_P\) ensures that the prover is a device in the \ac{PKI} of the 
\ac{CA}.

\Cref{DBSHW-overview} provides an overview.
For simplicity of the exposition, we present a version of \(\DBSHW\) that 
allows the user to run \(\PK[\alpha][A = g^\alpha]\) (the Schnorr 
identification scheme) with a distance bound.
This can, of course, be generalized to any \ac{ZKPK} for discrete logarithms, 
for instance, CL04~\cite{CLsignatures}.

@@

    \multicolumn{2}{l}{%
      Static \(\sk_P\in \ZZ_q, \sigma = (a, A, b, B, c)\)
    }
    & Static \(\pk = (\qCL, \gCL, \ggCL, e, \XCL, \YCL, \ZCL)\)
    \\

@@

    \(\xDH\rgets \ZZ_{\qDH}\)
    & \(\ProtoSendRight{\XDH = \gDH^{\xDH}}\)
    & \(\yDH\rgets \ZZ_{\qDH}\)
    \\

    \(k \gets \YDH^{\xDH}\)
    & \(\ProtoSendLeft{\YDH = \gDH^{\yDH}}\)
    & \(k \gets \XDH^{\yDH}\)
    \\

@@

    \(r_1, r_2\rgets \ZZ_{\qCL}\)
    &
    &
    \\

    \(\tilde a = a^{r_1},
      \tilde A = A^{r_1},
      \tilde b = b^{r_1},
      \tilde B = B^{r_1},
      \tilde c = (c^{r_1})^{r_2}\)
    & \(\ProtoSendRight{
        \tilde \sigma = (\tilde a, \tilde A, \tilde b, \tilde B, \tilde c)
      }\)
    &
    \\[0.5em]

    \(v_x = e(\XCL, \tilde a),
      v_{xy} = e(\XCL, \tilde b),
      V_{xy} = e(\XCL, \tilde B),
      v_s = e(\gCL, \tilde c)\)
    &
    & \(v_x = e(\XCL, \tilde a),
      v_{xy} = e(\XCL, \tilde b),
      V_{xy} = e(\XCL, \tilde B),
      v_s = e(\gCL, \tilde c)\)
    \\[0.5em]

    \(\rho\rgets \ZZ_q,
      \rhoDH \rgets \ZZ_{\qDH},
      \rhoCL_{\sk} \rgets \ZZ_{\qCL},
      \rhoCL_{r} \rgets \ZZ_{\qCL},
      \rhoCL_{r'} \rgets \ZZ_{\qCL}\)
    &
    &
    \\

    \(R\gets g^{\rho},
      \RDH\gets \gDH^{\rhoDH},
      \RCL\gets v_s^{\rhoCL_{r'}} v_{xy}^{\rhoCL_{\sk}} V_{xy}^{\rhoCL_r}\)
    & \(\ProtoSendRight{R, \RDH, \RCL}\)
    &
    \\

    \(\forall i\in \{1, \dotsc, m\}\):
    & \(\ProtoSendLeft{c_1, \dotsc, c_k}\)
    & \(c_1\rgets \bin^l, \dotsc, c_m\rgets \bin^l\)
    \\

    \(s_i\gets \rho - c_i\alpha \pmod q\)
    &
    &
    \\

    \(\sDH_i\gets \rhoDH - c_i \xDH \pmod \qDH\)
    &
    &
    \\

    \(\sCL_i^{(r')}\gets \rhoCL_{r'} + c_i r' \pmod \qCL\)
    &
    &
    \\

    \(\sCL_i^{(\sk)}\gets \rhoCL_{\sk} - c_i \sk \pmod \qCL\)
    &
    &
    \\

    \(\sCL_i^{(r)}\gets \rhoCL_{r} - c_i r \pmod \qCL\)
    &
    &
    \\

@@

    Fetch \(s_i, \sDH_i, \sCL_i^{(r')}, \sCL_i^{(\sk)}, \sCL_i^{(r)}\)
    & \(\ProtoSendRight{
      s_i, \sDH_i, \sCL_i^{(r')}, \sCL_i^{(\sk)}, \sCL_i^{(r)}}\)
    & Stop clock, record \(\Delta t\)
    \\

@@

    % null
    &
    & Accept if \(\Delta t < t_{\max}\) and
    \\

    % null
    &
    & \(R = A^{c_i} g^{s_i}\),
    \\

    % null
    &
    & \(\RDH = \XDH^{c_i} \gDH^{\sDH_i}\),
    \\

    % null
    &
    & \(v_x^{c_i} \RCL =
          v_s^{\sCL_i^{(r')}} v_{xy}^{\sCL_i^{(\sk)}} V_{xy}^{\sCL_i^{(r)}}\),
    \\

    %
    &
    & \(e(\tilde a, \ZCL) = e(\gCL, \tilde A)\),
    \\

    %
    &
    & \(e(\tilde a, \YCL) = e(\gCL, \tilde b)\),
    \\

    %
    &
    & \(e(\tilde A, \YCL) = e(\gCL, \tilde B)\).
    \\

  \end{tabular}
  \caption{%
    One-round protocol instance of the \(\DBSHW\) protocol instantiating 
    \(\PK[\alpha][A = g^\alpha]\).
    Each transmission (arrow in the diagram) uses \(\UWBPR\) (except in the 
    pre-setup phase).
    The protocol actually proves \(
      \PK[\alpha, \xDH, \sk, \sigma][
        A = g^\alpha \land
        \XDH = \gDH^{\xDH} \land
        \Blind[\sigma] = \Sign[_{\sk}][\sk_P]
      ].
    \)
    This is ties \(\xDH\), and consequently \(k\), to \(\sk_P\) and \(\alpha\).
    The protocol (setup, distance-bounding and verification phases) should be 
    repeated \(n\) times to achieve the desired soundness and distance-bounding 
    errors.
  }%
  \label{DBSHW-overview}
\end{figure*}

@@

We have public parameters \(\qCL, \gCL\in\GCL, \ggCL\in\GGCL, e\); where 
\(e\colon \GCL\times \GCL\to \GGCL\) is a bilinear map, \(\gCL\in \GCL\) and 
\(\ggCL\in \GGCL\) are generators of prime order \(\qCL\) (hence \(e(\gCL, 
\gCL) = \ggCL\)).

We let the \acg{CA} signing key be \[
  \sk = (\xCL, \yCL, \zCL)\rgets \ZZ_{\qCL}^3
\] and the public key be \[
  \pk = (\qCL, \gCL, \ggCL, e, \XCL, \YCL, \ZCL),
\] where \[
  \XCL = \gCL^{\xCL}, \quad
  \YCL = \gCL^{\yCL}, \quad
  \ZCL = \gCL^{\zCL}.
\]

The prover chooses \(\sk_P\rgets \ZZ_{\qCL}\) and computes a commitment \(M = 
\gCL^{\sk_P} Z^r\), where \(r\rgets \ZZ_{\qCL}\).
Now the prover convinces the \ac{CA} that he knows the key by running the 
following \ac{ZKPK} protocol: \(
  \PK[\sk_P, r][M = \gCL^{\sk_P} Z^r].
\)

Now the \ac{CA} computes the signature~\(\sigma = (a, A, b, B, c)\), where \[
  \alpha\rgets \ZZ_{\qCL}, \quad a\gets \gCL^\alpha, \quad
  A\gets \gCL^{\zCL}, \quad
  b\gets a^{\yCL}, \quad
  B\gets A^{\yCL}, \quad
  c\gets a^{\xCL} M^{\alpha \xCL \yCL}
\] and gives to the prover.

@@

To establish \(k\), we use the \ac{DHKE}.
The prover chooses \(\xDH \rgets \ZZ_{\qDH}\), the verifier chooses 
\(\yDH\rgets \ZZ_{\qDH}\) and they exchange \(\XDH = \gDH^{\xDH}, \YDH = 
\gDH^{\yDH}\) and both compute
\(k = \gDH^{\xDH \yDH}\), for some generator \(\gDH\) in \iac{DH} group of 
prime order \(\qDH\).

@@

The setup phase is similar to \(\DBS\) (\cref{SchnorrFigure}).
They use the agreed upon key~\(k = \gDH^{\xDH \yDH}\) as the key for 
\(\UWBPR\), instead of a static \(k\).
\(\DBS\) performs \(
  \PK[\alpha][A = g^\alpha]
\) (with a distance bound) whereas \(\DBSHW\) must perform \[
  \PK[\alpha, \xDH, \sk_P, \sigma][
    A = g^\alpha \land
    \XDH = \gDH^{\xDH} \land
    \Blind[\sigma] = \Sign[_{\sk}][\sk_P]
  ].
\] This means that the prover (and verifier) must prepare for the other proofs 
too.

The prover blinds the signature~\(\sigma\):
Choose \(r_1, r_2\rgets \ZZ_{\qCL}\) and compute \(\tilde \sigma = (\tilde a, 
\tilde A, \tilde b, \tilde B, \tilde c)\), where \[
  \tilde a = a^{r_1}, \quad
  \tilde A = A^{r_1}, \quad
  \tilde b = b^{r_1}, \quad
  \tilde B = B^{r_1}, \quad
  \tilde c = (c^{r_1})^{r_2}.
\]
The prover sends \(\tilde \sigma\) to the verifier.
Now, the prover and verifier each compute \[
  v_x\gets e(\XCL, \tilde a), \quad
  v_{xy}\gets e(\XCL, \tilde b), \quad
  V_{xy}\gets e(\XCL, \tilde B), \quad
  v_s\gets e(\gCL, \tilde c).
\]

The prover chooses \[
  \rho\rgets \ZZ_q, \quad
  \rhoDH\rgets \ZZ_{\qDH}, \quad
  \rhoCL_{\sk}\rgets \ZZ_{\qCL}, \quad
  \rhoCL_r\rgets \ZZ_{\qCL}, \quad
  \rhoCL_{r'}\rgets \ZZ_{\qCL}
\] computes \[
  R\gets g^\rho, \quad
  \RDH\gets \gDH^{\rhoDH}, \quad
  \RCL\gets v_s^{\rhoCL_{r'}} v_{xy}^{\rhoCL_{\sk}} V_{xy}^{\rhoCL_r}
\] and sends \(R, \RDH, \RCL\) to the verifier.

The verifier chooses \(m\) challenges of length \(l\), \(c_1\rgets \bin^l, 
\dotsc, c_m\rgets \bin^l\), and sends these to the prover.
The prover computes responses
\begin{multline*}
  s_i\gets \rho - c_i\alpha, \quad
  \sDH_i\gets \rhoDH - c_i \xDH, \\%\quad
  \sCL_i^{(r')}\gets \rhoCL_{r'} + c_i r', \quad
  \sCL_i^{(\sk)}\gets \rhoCL_{\sk} - c_i \sk, \quad
  \sCL_i^{(r)}\gets \rhoCL_r - c_i r
\end{multline*}
for \(i\in \{1, \dotsc, m\}\), where \(r' = r_2^{-1}\).

@@

The verifier decides which of the \(m\) challenges to use, \(i\rgets \{1, 
\dotsc, m\}\), sends its decision~\(i\) to the prover and starts its clock.
The prover instantly replies with the pre-computed \(s_i, \sDH_i, 
\sCL_i^{(r')}, \sCL_i^{(\sk)}, \sCL_i^{(r')}\).
Note that these values must be sent as a concatenation in \emph{one} \(\UWBPR\) 
reply.
The verifier stops the clock and records the round-trip time~\(\Delta t\).

@@

The verifier checks that \[
  R = A^{c_i} g^{s_i}, \quad
  \RDH = \XDH^{c_i} g^{\sDH_i}, \quad
  v_x^{c_i} \RCL = v_s^{s_{r'}} v_{xy}^{s_{\sk_P}} V_{xy}^{s_r}
\] and that the round-trip time \(\Delta t < t_{\max}\) does not exceed the 
time-limit~\(t_{\max}\) (corresponding to the desired distance bound).
Finally, the verifier checks the validity of \(\tilde \sigma\): \[
  e(\tilde a, \ZCL) = e(\gCL, \tilde A), \quad
  e(\tilde a, \YCL) = e(\gCL, \tilde b), \quad
  e(\tilde A, \YCL) = e(\gCL, \tilde B).
\]
